Supplement to Manuscript for MADA Project

### 2023-04-21 code/analysis_code/exploratory_analysis/exploratory_analysis.Rmd
Project Update

```{r, warning = FALSE, message = FALSE}
# Loading required packages
library(readr)
library(tidyverse)
library(dplyr)
library(survey)
library(naniar)
library(gmodels) #for ch-sq frequency table
library(vcd) # for three way frequency table
library(gtsummary)
library(openxlsx)
library(ggplot2)
library(ggthemes)
library(here)
```

@@

```{r}
# Recoding variables to work in making table
data_clean_maint <- data_clean_main
data_clean_maint$hispanic_status <- recode_factor(data_clean_maint$hispanic_status, '1' = 'Non-hispanic Female', '2' = 'Hispanic Female')
data_clean_maint$mammogram_status <- recode_factor(data_clean_maint$mammogram_status, '0' = 'No', '1' = 'Yes')
data_clean_maint$age <- recode_factor(data_clean_maint$age, '1' = '40-44 years', '2' = '45-49 years', '3' = '50-54 years', '4' = '55-59 years', '5' = '60-64 years', '6' = '65-69 years', '7' = '70-74 years')
data_clean_maint$income <- recode_factor(data_clean_maint$income, '1' = 'At or above poverty threshold', '2' = 'Below poverty threshold')
data_clean_maint$education <- recode_factor(data_clean_maint$education, '1' = 'Never attended/kindergarten only', '2' = 'Less than high school', '3' = 'High school', '4' = 'Some college or Associate degree', '5' = 'Bachelor degree and higher')
data_clean_maint$marital_status <- recode_factor(data_clean_maint$marital_status, '0' = 'Never married', '1' = 'Married', '2' = 'Single (separated/widowed/divorced)', '3' = 'Living with partner')
data_clean_maint$region_residence <- recode_factor(data_clean_maint$region_residence, '1' = 'Northeast', '2' = 'North Central/Midwest', '3' = 'South', '4' = 'West')
data_clean_maint$insurance_status <- recode_factor(data_clean_maint$insurance_status, '0' = 'No Coverage', '1' = 'Coverage')
data_clean_maint$health_status <- recode_factor(data_clean_maint$health_status, '1' = 'Excellent/very good/good', '2' = 'Fair/poor')
data_clean_maint$usual_medicalcare_status <- recode_factor(data_clean_maint$usual_medicalcare_status, '0' = 'No place', '1' = 'Usual place')
data_clean_maint$smoking_status <- recode_factor(data_clean_maint$smoking_status, '0' = 'Never', '1' = 'Current', '2' = 'Former')
data_clean_maint$alcohol_status <- recode_factor(data_clean_maint$alcohol_status, '0' = 'Never', '1' = 'Current', '2' = 'Former')
data_clean_maint$diabetes_status <- recode_factor(data_clean_maint$diabetes_status, '0' = 'No', '1' = 'Yes')

@@

# Lets create a single complete demographic characteristics table by year and hispanic_status of women
table1 <- data_clean_maint %>% 
    select(mammogram_status, age, income, education, marital_status, region_residence, insurance_status, health_status, usual_medicalcare_status, smoking_status, alcohol_status, diabetes_status, year, hispanic_status) %>%
    tbl_strata(
      strata = year,
      .tbl_fun = 
        ~ .x %>%
        tbl_summary(
        missing = "no",
        by = c("hispanic_status"),
        statistic = all_categorical() ~ "{n} ({p}%)"))
table1

# Lets save the above generated table as a `demographic_table1.rds` table in results folder
# Commenting this out so that it does not get created many times
# saveRDS(table1, file = (here("results", "exploratory_analysis_result", "demographic_table1.rds")))
```

@@

```{r}
# Summarizing mammogram_status and Hispanic_status in the given years (Type I)
data_clean_main1 <- data_clean_main
data_clean_main1$hispanic_status <- recode_factor(data_clean_main1$hispanic_status, '1' = 'Non-hispanic Female', '2' = 'Hispanic Female')

# Lets save this figure
# png("figure1_mammogram_ses.png", width = 10, height = 10, units = "in", res = 300)
ggplot(data = data_clean_main1, aes(x = mammogram_status)) +
  geom_histogram(fill = "white", colour = "red", stat = "count", position = position_dodge()) +
  facet_grid(hispanic_status ~ year) +
  scale_x_discrete(labels = c("No", "Yes")) +
  theme_bw()
# dev.off()

# Summarizing mammogram_status and Hispanic_status in the given years (Type II)
# png("figure1optional_mammogram_ses.png", width = 10, height = 10, units = "in", res = 300)
ggplot(data_clean_main1, aes(x = mammogram_status))+
  geom_bar(aes(fill = hispanic_status), stat = "count", color = "white",
    position = position_dodge(0.9)) +
  facet_wrap(~year) +
  scale_x_discrete(labels = c("No", "Yes"))
# dev.off()
```

>>>>

```{r, warning = FALSE, message = FALSE}
# Loading required packages
library(readr)
library(tidyverse)
library(dplyr)
library(survey)
library(naniar)
library(gmodels) #for chi-sq frequency table
library(vcd) # for three way frequency table
library(gtsummary)
library(openxlsx)
library(ggplot2)
library(ggthemes)
library(here)
```

@@

```{r}
# Recoding variables to work in making table
data_clean_maint <- data_clean_main
data_clean_maint$hispanic_status <- recode_factor(data_clean_maint$hispanic_status, '1' = 'NH', '2' = 'H')
data_clean_maint$mammogram_status <- recode_factor(data_clean_maint$mammogram_status, '0' = 'No', '1' = 'Yes')
data_clean_maint$age <- recode_factor(data_clean_maint$age, '1' = '40-44 years', '2' = '45-49 years', '3' = '50-54 years', '4' = '55-59 years', '5' = '60-64 years', '6' = '65-69 years', '7' = '70-74 years')
data_clean_maint$income <- recode_factor(data_clean_maint$income, '1' = 'At or above poverty threshold', '2' = 'Below poverty threshold')
data_clean_maint$education <- recode_factor(data_clean_maint$education, '1' = 'Never attended/kindergarten only', '2' = 'Less than high school', '3' = 'High school', '4' = 'Some college or Associate degree', '5' = 'Bachelor degree and higher')
data_clean_maint$marital_status <- recode_factor(data_clean_maint$marital_status, '0' = 'Never married', '1' = 'Married', '2' = 'Single (separated/widowed/divorced)', '3' = 'Living with partner')
data_clean_maint$region_residence <- recode_factor(data_clean_maint$region_residence, '1' = 'Northeast', '2' = 'North Central/Midwest', '3' = 'South', '4' = 'West')
data_clean_maint$insurance_status <- recode_factor(data_clean_maint$insurance_status, '0' = 'No Coverage', '1' = 'Coverage')
data_clean_maint$health_status <- recode_factor(data_clean_maint$health_status, '1' = 'Excellent/very good/good', '2' = 'Fair/poor')
data_clean_maint$usual_medicalcare_status <- recode_factor(data_clean_maint$usual_medicalcare_status, '0' = 'No place', '1' = 'Usual place')
data_clean_maint$smoking_status <- recode_factor(data_clean_maint$smoking_status, '0' = 'Never', '1' = 'Current', '2' = 'Former')
data_clean_maint$alcohol_status <- recode_factor(data_clean_maint$alcohol_status, '0' = 'Never', '1' = 'Current', '2' = 'Former')
data_clean_maint$diabetes_status <- recode_factor(data_clean_maint$diabetes_status, '0' = 'No', '1' = 'Yes')

@@

# Lets create a single complete demographic characteristics table by year and hispanic_status of women
table1 <- data_clean_maint %>% 
    select(mammogram_status, age, income, education, marital_status, region_residence, insurance_status, health_status, usual_medicalcare_status, smoking_status, alcohol_status, diabetes_status, year, hispanic_status) %>%
    tbl_strata(
      strata = year,
      .tbl_fun = 
        ~ .x %>%
        tbl_summary(
        missing = "no",
        by = c("hispanic_status"),
        statistic = all_categorical() ~ "{n} ({p}%"))
table1

# Saving the table
saveRDS(table1, file = (here("results", "exploratory_analysis_result", "demographic_table1.rds")))
```

@@

```{r}
# Recoding hispanic_status for the figure
data_clean_main1 <- data_clean_main
data_clean_main1$hispanic_status <- recode_factor(data_clean_main1$hispanic_status, '1' = 'Non-hispanic Female', '2' = 'Hispanic Female')

# Summarizing mammogram_status and Hispanic_status in the given years
figure1 <- ggplot(data_clean_main1, aes(x = mammogram_status))+
  geom_bar(aes(fill = hispanic_status), stat = "count", color = "white",
    position = position_dodge(0.9)) +
  facet_wrap(~year) +
  scale_x_discrete(labels = c("No", "Yes")) +
  scale_y_continuous(breaks = c(1000, 2000, 3000, 4000, 5000, 6000))
figure1

# Saving this figure
ggsave(filename = here("results", "exploratory_analysis_result", "figure1_mammogram_ses.png"), plot = figure1)
```

### 2023-04-21 code/analysis_code/statistical_analysis/1_model_fitting/model_fitting.Rmd
Project Update

```{r, warning = FALSE, message = FALSE}
# Loading required packages
library(readr)
library(tidyverse)
library(dplyr)
library(gtsummary)
library(ggplot2)
library(ggthemes)
library(here)
library(tidymodels)
library(performance)
library(vip)
```

@@

```{r}
# Comparing all the above models on the basis of model performance
model_performance <- compare_performance(log_fit_year, log_fit_hispanicstatus, log_fit_age, log_fit_income, log_fit_education, log_fit_maritalstatus, log_fit_regionresidence, log_fit_insurancestatus, log_fit_healthstatus, log_fit_usualmedicalcarestatus, log_fit_smokingstatus, log_fit_alcoholstatus, log_fit_diabetesstatus, log_fit_fullmodel, log_fit_year_age_hispanic, log_fit_demographic, log_fit_medical, log_fit_smoking_alcohol)
table7 <- model_performance

>>>>

```{r, warning = FALSE, message = FALSE}
# Loading required packages
library(readr)
library(tidyverse)
library(dplyr)
library(gtsummary)
library(ggplot2)
library(ggthemes)
library(here)
library(tidymodels)
library(performance)
library(vip)
library(gt)
```

@@

```{r}
# Comparing all the above models on the basis of model performance
model_performance <- compare_performance(log_fit_year, log_fit_hispanicstatus, log_fit_age, log_fit_income, log_fit_education, log_fit_maritalstatus, log_fit_regionresidence, log_fit_insurancestatus, log_fit_healthstatus, log_fit_usualmedicalcarestatus, log_fit_smokingstatus, log_fit_alcoholstatus, log_fit_diabetesstatus, log_fit_fullmodel, log_fit_year_age_hispanic, log_fit_demographic, log_fit_medical, log_fit_smoking_alcohol)
table7 <- model_performance
table7

### 2023-04-21 code/analysis_code/statistical_analysis/2_model_evaluation/model_evaluation.Rmd
Project Update

# Creating ROC curve and piping to the autoplot() method
data_clean_main_aug %>% 
  roc_curve(truth = mammogram_status, .pred_0) %>% 
  autoplot()

# Using roc_auc() for estimating the area under the curve 
data_clean_main_aug %>% 
  roc_auc(truth = mammogram_status, .pred_0)
```

@@

**ALTERNATIVE MODEL: USING TRAINED WORKFLOW TO PREDICT**

@@

# Creating ROC curve and piping to the autoplot() method
data_clean_main_aug2 %>% 
  roc_curve(truth = mammogram_status, .pred_0) %>% 
  autoplot()

>>>>

# Creating ROC curve and piping to the autoplot() method
roc_fullmodel <- data_clean_main_aug %>% 
  roc_curve(truth = mammogram_status, .pred_0) %>% 
  autoplot()
roc_fullmodel

# Saving the 'roc_fullmodel' figure
ggsave(filename = here("results", "statistical_analysis_result", "2_model_evaluation_result", "figure2_roc_fullmodel.png"), plot = roc_fullmodel)

# Using roc_auc() for estimating the area under the curve 
data_clean_main_aug %>% 
  roc_auc(truth = mammogram_status, .pred_0)

```

@@

**ALTERNATIVE MODEL: USING TRAINED WORKFLOW FOR PREDICTION**

@@

# Creating ROC curve and piping to the autoplot() method
roc_altmodel <- data_clean_main_aug2 %>% 
  roc_curve(truth = mammogram_status, .pred_0) %>% 
  autoplot()
roc_altmodel

# Saving the 'roc_altmodel' figure
ggsave(filename = here("results", "statistical_analysis_result", "2_model_evaluation_result", "figure3_roc_altmodel.png"), plot = roc_altmodel)

### 2023-04-21 code/analysis_code/statistical_analysis/3_machine_learning/machine_learning.Rmd
Project Update

@@

```{r}
# Fitting the final lasso model on split data
final_lasso_test <- final_lasso_wf %>% last_fit(data_split) 
final_lasso_test %>% collect_metrics()
```

>>>>

```{r}
# Let's collect the ROC AUC metric values for different models
final_lasso_test <- final_lasso_wf %>% last_fit(data_split) 
final_lasso_test %>% collect_metrics()
```

@@

```{r}
# Fitting the final lasso model on split data
ml_results <- data.frame(Model = c("Null", "Decision Tree", "Lasso"),
                      ROC_AUC = c(0.5, 0.7, 0.8))
ml_results

# Saving the above data frame
saveRDS(ml_results, file = (here("results", "statistical_analysis_result", "3_machine_learning_result", "machinelearning_results.rds")))
```

### 2023-04-21 products/manuscript/Manuscript.qmd
Project Update

# Summary/Abstract

@@

# Introduction

@@

# Methods

The study used the data from the National Health Interview Survey (NHIS). The NHIS survey collects information related to health, health care access, health behaviors of the civilian and non-institutionalized U.S. population and makes digital data files available for years 1963 to present (@blewett_lynn_a._ipums_2016). Users can create their own costumed NHIS data extracts for analysis from the NHIS website. For this study, data with the variables of interest for years: 2000, 2005, 2010, and 2015 were directly extracted from the NHIS website. Those specific 4 years were selected for the study as there were most data points available for those years and 5-years difference comparison could give great insights at the trends of mammogram screening over years. The study was restricted among Hispanic and non-Hispanic women aged 40-74 years as per the mammography screening recommendation by USPSTF and those age groups women fall under high-risk groups for developing breast cancer.

@@

The study was limited within 40-74 years Hispanic and non-Hispanic women with no report of cancer during the years 2000, 2005, 2010, and 2015. Exclusion of missing data points in 1 or more predictor variables yielded a final sample of 24848 subjects. \[\*\*References and Statistical analysis part to be added\]

# Results

## Exploratory/Descriptive analysis

@@

```{r, warning = FALSE, message = FALSE}
#| echo: FALSE
library(here)
library(knitr)

# Loading the `demographic_table1.rds` table which was created during the data exploratory process
result_table1 <- readRDS(here("results", "exploratory_analysis_result", "demographic_table1.rds"))
result_table1
```

@@

## Statistical analysis

**Table 2: Fitting full model using mammogram_status as outcome variable and other variables as predictor variables.**

@@

# Loading the `full_model.rds` table which was created during the statistical analysis process
result_table2 <- readRDS(here("results", "statistical_analysis_result", "1_model_fitting_result", "full_model.rds"))
print(result_table2, n = Inf, width = Inf)
```

# Discussion

@@

# Conclusions

@@

# References

>>>>

# 1. Summary/Abstract

@@

# 2. Introduction

@@

# 3. Methods

The study used the data from the National Health Interview Survey (NHIS). The NHIS survey collects information related to health, health care access, health behaviors of the civilian and non-institutionalized US population and makes digital data files available for years 1963 to present (@blewett_lynn_a._ipums_2016). Users can create their own costumed NHIS data extracts for analysis from the NHIS website. For this study, data with the variables of interest for years: 2000, 2005, 2010, and 2015 were directly extracted from the NHIS website. Those specific 4 years were selected for the study as there were most data points available for those years and 5-years difference comparison could give great insights at the trends of mammogram screening over years. The study was restricted among Hispanic and non-Hispanic women aged 40-74 years as per the mammography screening recommendation by USPSTF and those age groups women fall under high-risk groups for developing breast cancer.

@@

Data Cleaning:

Analysis:

The study was limited within 40-74 years Hispanic and non-Hispanic women with no report of cancer during the years 2000, 2005, 2010, and 2015. Exclusion of missing data points in 1 or more predictor variables yielded a final sample of 24848 subjects.

[** Statistical analysis part to be added]

# 4. Results

# 4.1 Exploratory/Descriptive analysis

@@

```{r warning = FALSE, message = FALSE}
#| echo: FALSE
library(here)
library(knitr)

# Loading the `demographic_table1.rds` table which was created during the data exploratory process

result_table1 <- readRDS(here("results", "exploratory_analysis_result", "demographic_table1.rds"))
result_table1 
```

@@

# 4.2 Statistical Analysis

# 4.2.1 Model Fitting

**Table 2: Full model fitting using mammogram_status as outcome variable and other variables as predictor variables.**

@@

# Loading the `full_model.rds` table which was created during the model fitting analysis process
result_table2 <- readRDS(here("results", "statistical_analysis_result", "1_model_fitting_result", "full_model.rds"))
print(result_table2, n = Inf, width = Inf)
```
**Table 3: Model Performance Comparisons.**

# Loading the `model_performance_comparison.rds` table which was created during the model fitting analysis process
result_table2 <- readRDS(here("results", "statistical_analysis_result", "1_model_fitting_result", "model_performance_comparison.rds"))
print(result_table2)
```

# 4.2.2 Model Evaluation

**Figure 2: ROC Curve for the full model**

```{r}
#| label: Figure 2
#| fig-cap: "ROC Curve for the full model."
#| echo: FALSE

# Loading the `figure2_roc_fullmodel.png` figure which was created during the model evaluation process
include_graphics(here("results", "statistical_analysis_result", "2_model_evaluation_result", "figure2_roc_fullmodel.png"))
```

**Figure 3: ROC Curve for the alternative model**

```{r}
#| label: Figure 3
#| fig-cap: "ROC Curve for the alternative model."
#| echo: FALSE

# Loading the `figure3_roc_altmodel.png` figure which was created during the model evaluation process
include_graphics(here("results", "statistical_analysis_result", "2_model_evaluation_result", "figure3_roc_altmodel.png"))
```

# 4.2.3 Machine Learning

**Table 4: Model Selection using Null, Decision tree and Lasso methods.**

# Loading the `machinelearning_results.rds` table which was created during the machine learning analysis process
result_table3 <- readRDS(here("results", "statistical_analysis_result", "3_machine_learning_result", "machinelearning_results.rds"))
print(result_table3)
```

# 5 Discussion

@@

# 6 Conclusions

@@

# 7 References

### 2023-04-21 products/supplement/Supplementary_Material.qmd
Project Update

```{r}
#| label: Additional Figure 1
#| fig-cap: "Decision tree model."
#| echo: FALSE
knitr::include_graphics("/results/statistical_analysis_result/3_machine_learning_result/figure2_decision_tree_model.png")
```

```{r}
#| label: Additional Figure 2
#| fig-cap: "Lasso model."
#| echo: FALSE
knitr::include_graphics("/results/statistical_analysis_result/3_machine_learning_result/figure3_lasso_model.png")
```

>>>>

```{r}
#| label: Additional Figure 1
#| fig-cap: "Lasso model."
#| echo: FALSE
knitr::include_graphics("/results/statistical_analysis_result/3_machine_learning_result/figure3_lasso_model.png")
```
